---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ntdr: Retrieve data from the National Transit Database

<!-- badges: start -->
<!-- badges: end -->

The `ntdr` package provides a convenient interface to retrieve monthly data from the National Transit Database (NTD). Without this package, getting NTD data involves going to the NTD website and downloading and processing a complex Excel workbook.

## Installation

You can install the development version of ntdr from Github:

``` r
devtools::install_github()
```

## Example

Basic usage:

```{r example}
library(ntdr)
ntd_madison <- get_ntd(data_type = "adjusted", ntd_variable = "UPT", agency = "City of Madison", modes = "MB")
head(ntd_madison)
```

The data are returned in a long format, which makes it easy to create plots and do other cool things:

```{r ridership-chart}
library(ggplot2)
get_ntd(agency = c("City of Madison", "Capital Area Transportation Authority"), modes = "MB") |> 
  dplyr::filter(TOS == "DO") |> 
  ggplot(aes(date, value, color = Agency)) +
  geom_line() +
  labs(title = "Monthly unlinked passenger trips in Madison and Lansing")
```

This package is still heavily under development and not ready for production use. NTD data has many quirks, and so please use it with care.
