---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ntdr: Retrieve data from the National Transit Database

<!-- badges: start -->
<!-- badges: end -->

The `ntdr` package provides a convenient interface to retrieve monthly data from the National Transit Database (NTD). Without this package, getting NTD data involves going to the NTD website and downloading and processing a complex Excel workbook.

This package was mainly written for my own use, is still under development and NTD data has many quirks. Please use it with care and feel report to [report any issues](https://github.com/vgXhc/ntdr/issues). 

To learn more about the package, visit the package website: [https://vgxhc.github.io/ntdr/](https://vgxhc.github.io/ntdr/). 

## Installation

You can install the development version of ntdr from Github:

``` r
remotes::install_github("https://github.com/vgXhc/ntdr")
```

## Example

Basic usage:

```{r example}
library(ntdr)
ntd_madison <- get_ntd(data_type = "adjusted", ntd_variable = "UPT", agency = "City of Madison", modes = "MB")
head(ntd_madison)
```
