---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ntdr: Retrieve data from the National Transit Database

<!-- badges: start -->
[![R-CMD-check](https://github.com/vgXhc/ntdr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/vgXhc/ntdr/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/ntdr)](https://cran.r-project.org/package=ntdr)
<!-- badges: end -->

The `ntdr` package provides a convenient interface to retrieve monthly data from the National Transit Database (NTD). Without this package, getting NTD data involves going to the NTD website and downloading and processing a complex Excel workbook.

The NTD data returned by this package has many quirks. Please use it with care and [report any issues](https://github.com/vgXhc/ntdr/issues). 

To learn more about the package, visit the package website: [https://vgxhc.github.io/ntdr/](https://vgxhc.github.io/ntdr/). 

## Installation

```{r, eval = FALSE}
# The latest stable version can be installed from CRAN:
install.packages("ntdr")


# Alternatively, install the latest development version from GitHub:
remotes::install_github("vgXhc/ntdr")
```



## Example

Basic usage:

```{r example}
library(ntdr)
ntd_madison <- get_ntd(data_type = "adjusted", ntd_variable = "UPT", agency = "City of Madison", modes = "MB")
head(ntd_madison)
```
