---
title: "ntdr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ntdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The ``ntdr`` package is an easy way to access [National Transit Database](https://www.transit.dot.gov/ntd/ntd-data) from R. The package is available on Github, and you can install it from there with `remotes::install_github()`

```{r eval=FALSE}
remotes::install_github("https://github.com/vgXhc/ntdr", build_vignettes = TRUE)
```

In addition to loading the `ntdr` package we also load the `tidyverse`.
```{r setup}
library(ntdr)
library(tidyverse)
```

# `get_ntd()`
`get_ntd()` is the main function of the package. It doesn't have any required parameters:

```{r bare-get-ntd-call}
ntd_data <- get_ntd()
ntd_data
colnames(ntd_data)
```

You see that the package downloads a fairly large `xlsx` file from the web and returns a tibble with `r nrow(ntd_data)` rows and `r ncol(ntd_data)` columns. The first two columns are identifiers for the transit agency; followed by a human-readable agency name. Note that the agency name may not be what you expect. For example, the name of our local agency in Madison (Wisconsin) is ["Metro Transit"](https://www.cityofmadison.com/metro). But in the NTD data it is listed as "City of Madison". So if you cannot find your agency, use the `uza_name` variable described below.

NTD data go back as far as 2002, and some agencies no longer actively report data, report them under a different ID, or don't even exist anymore. This is reflected in the `active` column. `reporter_type` most commonly takes on the `Full reporter` value, but especially smaller or rural systems may have a different value for this variable. For agencies that aren't full reporters, the NTD data may include projections rather than actually reported data. 

`uza` is an identifier for [urbanized areas](https://www.fhwa.dot.gov/planning/census_issues/urbanized_areas_and_mpo_tma/faq/page01.cfm#q2) and `uza_name` has the name of that area (this will usually be how you will find your local agency). 

`modes` denotes the type of transit reported on. 

```{r}
ntd_data |> 
  count(modes)
```

There are a lot of different modes, including rather obscure ones like "Inclined Plane" (`IP`) or "Alaska Railroad" (`AR`). You can find documentation of the different modes [here](https://ftis.org/iNTD-Urban/modes.pdf).

The `tos` variable represents the "type of service":

```{r}
ntd_data |> 
  count(tos, sort = T)
```

The most common values are `DO`, which is directly operated service, i.e. a transit agency running their own service; and `PT` for "purchased transportation", i.e. a transit agency contracting out services. Often agencies will have an entry for both of these, with `DO` being the regular, fixed route service and `PT` being paratransit or other more specialized forms of transit.

Finally, the `month` and `value` variables provide the actual transit data for a given month. If you call `get_ntd()` without any additional parameters, it will return the `unlinked passenger trips` (UPT) metric for all agencies, modes, and types of service.


# Plot the data
The data are returned in a long format, which makes it easy to create plots:

```{r ridership-chart}
get_ntd(agency = c("City of Madison", "Capital Area Transportation Authority"), modes = "MB") |> 
  dplyr::filter(tos == "DO") |> 
  ggplot(aes(month, value, color = agency)) +
  geom_line() +
  labs(title = "Monthly unlinked passenger trips in Madison and Lansing") +
  theme_minimal()
```



